\pkg{sdmTMB}: An \proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields

### 2023-11-28 doc/paper-jss/sdmTMB-paper.Rnw
add note on centring SVCs
use SVC shorthand in more places

We will fit counts using a negative binomial \citep[NB2,][]{hilbe2011} distribution, random intercepts for year, spatial and spatiotemporal random fields, and a spatially varying coefficient associated with the NAO.

@@

In addition to the output seen for other models, we now have a section for random intercepts and a standard deviation for the spatially varying coefficient random field.
Given our model specification, all random fields are sharing a single Matérn range.
We can also check the confidence intervals (CIs) on the main effect of NAO and see that they overlap zero.

@@

\subsection{Extracting the spatially varying coefficient effects with uncertainty}

The spatially varying coefficient effect at any point in space is the combination of the main effect and spatially varying random effect for \code{nao}.
Mean estimates of the spatially varying random effect are available in the output of \fct{predict.sdmTMB} in a column starting with \code{zeta\_s} (in this case would appear as \code{zeta\_s\_nao}).
However, we might wish to combine the fixed and random components of a spatially varying effect and assess the uncertainty of these combined predictions.
We illustrate a way of accomplishing this by simulating from the parameter precision matrix assuming the joint parameter covariance matrix is multivariate normal.
We do this by specifying a non-null number of simulation draws to \code{nsim} in \fct{predict.sdmTMB}.
By default, \code{nsim > 0} will return a matrix of draws for the overall prediction.
Here, we instead specify that we want to return draws for the \code{zeta\_s} ($\zeta_{\bm{s}}$) random field, which is the spatially varying random field.
This returns a matrix where each row matches a row of \code{newdata} and each column is a simulation draw.
We then use the function \fct{spread\_sims} to draw 200 simulations for the parameters themselves.
Because the simulations are stored in different dimensions for these two functions, the random field draws must be transposed \code{t()} before combining the main effect draws (\code{sims\$nao}) with the random field values \code{zeta\_s}.
Next, we can calculate the median, and upper and lower quantiles for each column of data, which correspond to the rows in the data provided.
For this example we use thresholds of 0.1 and 0.9 representing CIs of 80\%.

>>>>

We will fit counts using a negative binomial \citep[NB2,][]{hilbe2011} distribution, random intercepts for year, spatial and spatiotemporal random fields, and a SVC associated with the NAO.
Centring and scaling variables is especially helpful when estimating SVCs.
NAO is an index so it can be left as is, but including it as a main effect as well helps ensure that the SVCs will have a mean near 0.

@@

In addition to the output seen for other models, we now have a section for random intercepts and a standard deviation for the SVC random field.
Given our model specification, all random fields are sharing a single Matérn range.
We can also check the confidence intervals (CIs) on the main effect of NAO and see that they overlap zero.

@@

\subsection{Extracting the spatially varying effects with uncertainty}

The spatially varying effect at any point in space is the combination of the main effect and SVC random effect for \code{nao}.
Mean estimates of the SVC random effect are available in the output of \fct{predict.sdmTMB} in a column starting with \code{zeta\_s} (in this case would appear as \code{zeta\_s\_nao}).
However, we might wish to combine the fixed and random components of a spatially varying effect and assess the uncertainty of these combined predictions.
We illustrate a way of accomplishing this by simulating from the parameter precision matrix assuming the joint parameter covariance matrix is multivariate normal.
We do this by specifying a non-null number of simulation draws to \code{nsim} in \fct{predict.sdmTMB}.
By default, \code{nsim > 0} will return a matrix of draws for the overall prediction.
Here, we instead specify that we want to return draws for the \code{zeta\_s} ($\zeta_{\bm{s}}$) random field, which is the SVC random field.
This returns a matrix where each row matches a row of \code{newdata} and each column is a simulation draw.
We then use the function \fct{spread\_sims} to draw 200 simulations for the parameters themselves.
Because the simulations are stored in different dimensions for these two functions, the random field draws must be transposed \code{t()} before combining the main effect draws (\code{sims\$nao}) with the random field values \code{zeta\_s}.
Next, we can calculate the median, and upper and lower quantiles for each column of data, which correspond to the rows in the data provided.
For this example we use thresholds of 0.1 and 0.9 representing CIs of 80\%.
